Theoretical prediction of disrupted Min oscillation in flattened \emph{Escherichia coli}

### 2015-03-17 paper/plos.tex
more edits to paper Tested everything on crystal.

\fixme{
  First referee:
  \begin{enumerate}
  \item Add discussion of ``platform'' --- wrote our own code,
    publically available (cite github?).
  \item Make 3D geometries available? Cite that the publically
    available code can generate them?
  \item Claims differential equations aren't needed.
  \item Claims wild-type movie-ish plot not needed.
  \item Points out problems with Huang 2003:
    \begin{enumerate}
    \item \blue{David.  I've added some sentences discussing the
      starry night in the context of the membrane diffusion.  I didn't
      say anything about a one half period allowing diffusion to smear
      an entire pole, because it seems to contradict the second part
      of this paragraph.  What do you think?} Lacks diffusion on the
      membrane. Supported by Meacci 2006, which experimentally tries
      to measure diffusion on membrane, has a tough time, but shows
      that definitely diffuses.
    \item Underestimates diffusion in cytoplasm by a factor of 5-6.5.
      Supported by Meacci 2006, who concludes a rate of 16 microns
      squared per sec for MinD and 10 microns squared per sec for
      MinE.
    \item Uses MinD-ADP -> MinD-ATP that is 3 times lower than
      experimental bound. Halatek uses ADP->ATP of 1/6 s-1. They cite
      Meacci 2006, who state that there is a lower bound on reference
      time of 3 seconds, so this constant has an upper bound of 1/3
      s-1.
    \item (Still debated) No evidence that DE recruits to membrane.
    \item Works in only a limited part of parameters phase space, and
      could not account for important things like
      temperature-independence.
    \end{enumerate}
  \item Height should be bigger to compare with experiment.  0.4-0.6
    microns.  The 0.25 micron height of the cells used in this study
    does not represent the experimental settings well. The simulations
    are based on the flattened cell shapes reported from the
    experiments by Mannik et al 2009 and 2012. In these experimental
    studies, E. coli were shown to be able to grow in channels as low
    as 0.6 micron in hard silicon structures, and no more than 0.4
    micron (see Mannik 2009). Indeed, 0.25 micron channels were
    constructed with PDMS ceilings, which is a rather flexible polymer
    allowing bacteria to deform it to effectively higher
    channels. Thus the simulations would be more realistic if
    performed in the range of 0.4-0.6 micron thickness.

    \emph{If you can't make it grow in a .3 micron (rigid) silicon channel,
      then we and you observe it growing in a .25 micron flexible
      channel, it presumably must be able to grow because it deformed
      the channel to larger than .3 microns.}
  \item Starry night in the wild-type cells is a problem.
  \item Better explanation of correlation function (which is
    interesting!).
  \item Shape A is the interesting bit in our main figure.
  \item Would like more analysis of stochastic locations in different
    geometries.  Clusters?  Could add more geometries?
  \item Bold conclusions about flattening.  Talk about both width and
    thickness.  Compare Halatek 2012.  Wants systematically more
    geometries again.
  \item Is doubtful of our correlation plots relative to those of
    Fange and Elf 2006.
  \item Clarify (esp in conclusions) what we mean by flattening and
    irregular shape.  ``breaking the cylindrical symmetry and
    flattening the cells.'' versus ``breaking the mirror plane
    symmetry of the cross-section of a flattned cell.''
  \item[TODO suggestion] Switch to Halatek and Frey
  \item[TODO suggestion] Thoroughly explor shape parameter space
  \item[TODO suggestion] Check robustness of our conclusions with
    respect to changes in (reaction-diffusion?) parameters.
  \item Check order of figures and their references.
  \item Cite biological (non-computational) works e.g. Corbin 2008.
  \end{enumerate}
  Second referee:
  \begin{enumerate}
  \item Nitpicks about language of deterministic model versus
    stochastic model.  (Read carefully)
  \item Explain in greater detail the obviousness (maybe) that the
    deterministic approach doesn't have fluctuations.  RESPOND:
    chaotic systems do deterministically result in fluctuations.
    Explain this somewhere, and discuss triviality of the conclusion
    that deterministic is wrong because it does not exhibit these
    irregularities.
  \item Would like temporal averages over cell-cycle time scales.
  \item More discussion of how flattening leads to spatial
    irregularities.  Compare with large nearly spherical e. coli Shih
    2005.  See also Halatek and Frey 2012 for rotating oscillation
    axis due to destabilization of short modes.  How does this connect
    with our stability analysis?
  \item Should mention more recent revisions of the model.
  \item Fange and Elf do not use the same reaction-diffusion
    equations.
  \item Clarify statement about 1000 proteins and stochastic
    fluctuation.  Reword about a continuum description omitting
    information?
  \item Do not call Huang parameters ``wild type values.''  They are
    fitted parameters.
  \item Add methods section for numerics (deterministic).
  \item Add details of the linear stability analysis.
  \item (From jeff): still need to fix periods on correlation plots
  \end{enumerate}
}

\fixme{Redo stability analysis with larger thickness, whatever we pick.}

@@

A number of models of the MinD protein system have been developed that
accurately describe its basic oscillatory nature.
%
Early models involved free proteins that affect each others' rates of
diffusion and membrane attachment, but do not combine into compound
states~\cite{meinhardt2001pattern}.  In 2003 Huang improved upon this
approach with a simple and very successful simulation model based on
MinD-MinE combination, ATPase hydrolysis, and MinD membrane attachment
that exhibits accurate MinD oscillations in cylindrical
cells~\cite{huang2003dynamic}. In this model cytoplasmic MinD is
recruited to the membrane by MinD that is already clustered there
(following observed non-linear attachment of MinD on the cell
membrane~\cite{hu2002dynamic,shih2002division}), and is stationary
once attached.
%
Several models
\cite{fange2006noise,bonny2013membrane,halatek2012highly} modify
Huang's original model to more accurately treat experimental findings
published after his 2003 paper
\cite{meacci2006mobility,loose2011min}. These models introduce slow
diffusion along the membrane (two to three orders of magnitude slower
than that within the cytoplasm), revise the process of MinD
recruitment to the membrane, and change some of the constants used in
the model. The Bonny 2013 model assumes an instantaneous nucleotide
exchange rate, and allows MinE to remain independently attached to the
membrane after dissasociating with MinD~\cite{bonny2013membrane}.
%

@@

Our diffusion and reaction rates are shown below.  We are interested
primarily in the effect of cellular size and shape on the protein
oscillations, so we follow Huang\cite{huang2003dynamic} and do not
deviate from the wild-type values used in the cited work.

@@

Huang's simulations use total MinD and MinE concentrations of
$1,000/\micron$ and $350/\micron$, respectively, in a cylindrical cell
of radius $0.5\micron$, and in our (non-cylindrical) cells we use the
same number of proteins per unit volume.  These concentration values
are $1273\micron^{-3}$ and $446\micron^{-3}$, respectively. We have
chosen to write our own simulation platform and membrane creation
tools from scratch, instead of using existing available software used
in previous studies\fixme{cite previous studies} (we believe there is
value in re-creating results with alternative software).  Our
simulations take place within a three-dimensional Cartesian grid that
has a grid spacing of .05\micron, and our cell shapes are specified as
the zero of an analytic function which can generate the geometry
presented in this paper.  Our code is publically available on
github.com~\fixme{cite github account in references? Yes!}

@@

Huang \emph{et al.}~\cite{huang2003dynamic} has performed a linear
stability analysis on a cylindrical model which shows an upper limit
on a steady state solution of a $2\micron$ half wavelength.  Cells
with dimensions longer than this spontaneously develop spatial
oscillations, while cells that are shorter relax into a homogeneous
steady state.  We have performed a similar linear stability analysis
on an infinite slab with a thickness equal to that of our flattened
cells ($0.40\micron$).  We first solve the model's five differential
equations for the steady state solution, under the constraint that the
total number MinD and MinE molecules matches our simulation.  We write
down as a matrix the linear response of the time derivatives of the
density to small, spatially harmonic density perturbations around the
steady-state density.  The system is stable at this wavelength
provided all eigenvalues of the matrix are negative, indicating that
all perturbations with this wavelength will decay.  We find the
stability limit by searching for the largest wavelength at which the
system is stable.  Through this method we arrive at an upper half
wavelength stability limit of $1.56\micron$ for our $0.40\micron$
flattened cells. As expected, when decreasing the lengths and widths
of our simulated flattened cells so that the longest distance across
the cell is less than this length, the cells stop exhibiting any
oscillatory behavior.  The deterministic model relaxes into a
motionless state and the stochastic model exhibits random fluctuations
without spatial oscillations.

@@

\blue{(5a)} Both Fig.~\ref{image-p} and its associated movies (see
supplementary material), show that the stochastic model exhibits a
``starry night'' effect characterized by spatially fixed points of
protein density build up, as recruitment leads to clusters of MinD
forming on the wall and then subsequently dissipating.  We attribute
this effect to the fact that the Huang model does not include
diffusion along the membrane, which has been experimentally verified
yet found to be two orders of magnitude less than diffusion in the
cytoplasm~\cite{meacci2005min}. Even ignoring this effect, however, it
is clear from Fig.~\ref{image-p} that the stochastic simulation
results in a protein distribution that is less spatially regular than
the deterministic model predicts.  This naturally leads to
irregularity in the temporal periodicity of the system, which we
quantify by examining the temporal correlation function of the total
MinD found in two opposite polar regions.
%
This correlation function, which is displayed in Fig.~\ref{corr-pill},
is given by
\begin{equation}
  C(\tau) \propto \int
  (N_{\textit{top}}(t) - \bar N_{\textit{top}})
  (N_{\textit{bottom}}(t+\tau) - \bar N_{\textit{bottom}})dt
\end{equation}
where $N_{\textit{top}}(t)$ and $N_{\textit{bottom}}(t)$ are the total
MinD proteins in the top and bottom thirds of a cell.  These plots
help us in studying the periods and the regularity and stability of
oscillations.  These curves are well fit with a simple decoherence
model, given by the equation
\begin{equation}
  C(\tau) = -\cos\left(\frac{2\pi\tau}{T}\right) e^{-\frac{\tau}{\tau_c}}
\end{equation}
where $T$ is the period and $\tau_c$ is the coherence time.  In the
case of the stochastic model the pill exhibits coherence times of
about 8 periods and in the case of the deterministic model oscillate
indefinitely with complete coherence, indicating that its behavior is
perfectly periodic.

\begin{figure*}
  \centering
  \hspace{-3cm}
  \includegraphics[width=16cm]{../paper/plot-ave}
  \caption{We display here arrows depicting successive maxima in space
    and and time overlayed on a color plot of the total MinD density
    averaged over the same time period.  The simulation time covered
    for the flattened cells is 350 seconds, which is the same
    period of time depicted in the experimental data plots of Mannik
    \emph{et al.}~\cite{mannik2012robustness}.  For the wild-type pill
    shape, we only cover 250 seconds, in order to provide a
    useful comparison due to its shorter oscillation period.  The top
    row shows plots published by Mannik \emph{et al.} of the MinD maxima behavior
    and the bottom two rows show our simulations using the stochastic
    and deterministic models, respectively.  We simulated
    approximations to the two shapes observed by Mannik, which we call
    \emph{shape A} and \emph{shape B}.  In addition, we studied two
    flattened stadium shapes which we call \emph{stadium A} and
    \emph{stadium B} corresponding in aspect ratio and thickness to
    the two experimental shapes.  The spatial length scale of all the
    figures shown was identical. Each of the flattened cell shapes
    uses the same color scale for the number of proteins per unit
    area.  Finally, we display the natural pill shape, which was also
    featured in Figs.~\ref{image-p} and~\ref{corr-pill}, with a
    different color scale to reflect the thicker cell containing more
    proteins per cross-sectional area.  }
  \label{randst-plot-ave}
\end{figure*}

@@

In every case, including the wild-type pill shape, we see irregularity
in the location of the maxima when using the stochastic model.  The
deterministic model shows uniformly bipolar oscillation, with the
exception of shape \emph{B}, which develops weak intermediate maxima
on the right-hand edge of the cell.  From these results, we conclude
that the deterministic model is inadequate to explain experimental
observations of the locations of density maxima of the MinD protein.

@@

In contrast to the stochastic model, the original deterministic
version of the model~\cite{huang2003dynamic} predicts behavior that
contradicts experimental observations.  Specifically, the
deterministic model predicts robust and periodic bipolar oscillation
in irregularly shaped cells, with a minor intermediate maximum
appearing in shape \emph{B}.  Thus we conclude that this method is
inconsistent with experiment, and cannot be relied upon for
predictions of the behavior of the MinD system.  This is the most
clear example of the inadequacy of the deterministic simulation method
in reproducing experimental behavior of the MinD system to date.
Until now the results of deterministic and stochastic simulation have
largely coincided, with the stochastic method showing minor
differences in behavior when compared to deterministic
models~\cite{kerr2006division, fange2006noise, huang2004min,
  kruse2007experimentalist}.

>>>>

\fixme{
  First referee:
  \begin{enumerate}
  \item \blue{Done} Add discussion of ``platform'' --- wrote our own code,
    publically available (cite github?).
  \item \blue{Done} Make 3D geometries available? Cite that the publically
    available code can generate them?
  \item Claims differential equations aren't needed.
  \item Claims wild-type movie-ish plot not needed.
  \item Points out problems with Huang 2003:
    \begin{enumerate}
    \item \blue{David.  I've added some sentences discussing the
      starry night in the context of the membrane diffusion.  I didn't
      say anything about a one half period allowing diffusion to smear
      an entire pole, because it seems to contradict the second part
      of this paragraph.  What do you think?} Lacks diffusion on the
      membrane. Supported by Meacci 2006, which experimentally tries
      to measure diffusion on membrane, has a tough time, but shows
      that definitely diffuses.
    \item Underestimates diffusion in cytoplasm by a factor of 5-6.5.
      Supported by Meacci 2006, who concludes a rate of 16 microns
      squared per sec for MinD and 10 microns squared per sec for
      MinE.
    \item Uses MinD-ADP -> MinD-ATP that is 3 times lower than
      experimental bound. Halatek uses ADP->ATP of 1/6 s-1. They cite
      Meacci 2006, who state that there is a lower bound on reference
      time of 3 seconds, so this constant has an upper bound of 1/3
      s-1.
    \item (Still debated) No evidence that DE recruits to membrane.
    \item Works in only a limited part of parameters phase space, and
      could not account for important things like
      temperature-independence.
    \end{enumerate}
  \item \blue{Done} Height should be bigger to compare with experiment.  0.4-0.6
    microns.  The 0.25 micron height of the cells used in this study
    does not represent the experimental settings well. The simulations
    are based on the flattened cell shapes reported from the
    experiments by Mannik et al 2009 and 2012. In these experimental
    studies, E. coli were shown to be able to grow in channels as low
    as 0.6 micron in hard silicon structures, and no more than 0.4
    micron (see Mannik 2009). Indeed, 0.25 micron channels were
    constructed with PDMS ceilings, which is a rather flexible polymer
    allowing bacteria to deform it to effectively higher
    channels. Thus the simulations would be more realistic if
    performed in the range of 0.4-0.6 micron thickness.

    \emph{If you can't make it grow in a .3 micron (rigid) silicon channel,
      then we and you observe it growing in a .25 micron flexible
      channel, it presumably must be able to grow because it deformed
      the channel to larger than .3 microns.}
  \item \blue{David} Starry night in the wild-type cells is a problem.
  \item Better explanation of correlation function (which is
    interesting!).
  \item Shape A is the interesting bit in our main figure.
  \item Would like more analysis of stochastic locations in different
    geometries.  Clusters?  Could add more geometries?
  \item Bold conclusions about flattening.  Talk about both width and
    thickness.  Compare Halatek 2012.  Wants systematically more
    geometries again.
  \item Is doubtful of our correlation plots relative to those of
    Fange and Elf 2006.
  \item Clarify (esp in conclusions) what we mean by flattening and
    irregular shape.  ``breaking the cylindrical symmetry and
    flattening the cells.'' versus ``breaking the mirror plane
    symmetry of the cross-section of a flattned cell.''
  \item[TODO suggestion] Switch to Halatek and Frey
  \item[TODO suggestion] Thoroughly explor shape parameter space
  \item[TODO suggestion] Check robustness of our conclusions with
    respect to changes in (reaction-diffusion?) parameters.
  \item Check order of figures and their references.
  \item Cite biological (non-computational) works e.g. Corbin 2008.
  \end{enumerate}
  Second referee:
  \begin{enumerate}
  \item Nitpicks about language of deterministic model versus
    stochastic model.  (Read carefully)
  \item \blue{David} Explain in greater detail the obviousness (maybe) that the
    deterministic approach doesn't have fluctuations.  RESPOND:
    chaotic systems do deterministically result in fluctuations.
    Explain this somewhere, and discuss triviality of the conclusion
    that deterministic is wrong because it does not exhibit these
    irregularities.
  \item Would like temporal averages over cell-cycle time scales.
  \item More discussion of how flattening leads to spatial
    irregularities.  Compare with large nearly spherical e. coli Shih
    2005.  See also Halatek and Frey 2012 for rotating oscillation
    axis due to destabilization of short modes.  How does this connect
    with our stability analysis?
  \item \blue{David} Should mention more recent revisions of the model.
  \item Fange and Elf do not use the same reaction-diffusion
    equations.
  \item Clarify statement about 1000 proteins and stochastic
    fluctuation.  Reword about a continuum description omitting
    information?
  \item \blue{Done} Do not call Huang parameters ``wild type values.''  They are
    fitted parameters.
  \item Add methods section for numerics (deterministic).
  \item \blue{Done} Add details of the linear stability analysis.
  \item \blue{David} Still need to fix periods on.  \blue{About this
    point - I updated the data and it seems the result is defferent
    from before - the stadium shapes seem to be temporally correlated
    for much longer compared with the shape A than before.}
    correlation plots
  \end{enumerate}
}

@@

A number of models of the MinD protein system have been developed that
accurately describe its basic oscillatory nature.
%
Early models involved free proteins that affect each others' rates of
diffusion and membrane attachment, but do not combine into compound
states~\cite{meinhardt2001pattern}.  In 2003 Huang improved upon this
approach with a simple and very successful simulation model based on
MinD-MinE combination, ATPase hydrolysis, and MinD membrane attachment
that exhibits accurate MinD oscillations in cylindrical
cells~\cite{huang2003dynamic}. In this model cytoplasmic MinD is
recruited to the membrane by MinD that is already clustered there
(following observed non-linear attachment of MinD on the cell
membrane~\cite{hu2002dynamic,shih2002division}), and is stationary
once attached.
%
\blue{(5)} Several models
\cite{fange2006noise,bonny2013membrane,halatek2012highly} modify
Huang's original model to more accurately treat experimental findings
published after his 2003 paper
\cite{meacci2006mobility,loose2011min}. These models introduce slow
diffusion along the membrane (two to three orders of magnitude slower
than that within the cytoplasm), revise the process of MinD
recruitment to the membrane, and change some of the constants used in
the model. The Bonny 2013 model assumes an instantaneous nucleotide
exchange rate, and allows MinE to remain independently attached to the
membrane after dissasociating with MinD~\cite{bonny2013membrane}.
%

@@

Our diffusion and reaction rates are shown below.  We are interested
primarily in the effect of cellular size and shape on the protein
oscillations, so we follow Huang\cite{huang2003dynamic} and do not
deviate from the \blue{(8)} parameter values used in the cited work.

@@

Huang's simulations use total MinD and MinE concentrations of
$1,000/\micron$ and $350/\micron$, respectively, in a cylindrical cell
of radius $0.5\micron$, and in our (non-cylindrical) cells we use the
same number of proteins per unit volume.  These concentration values
are $1273\micron^{-3}$ and $446\micron^{-3}$,
respectively. \blue{(1,2)} We have chosen to write our own simulation
platform and membrane creation tools from scratch, instead of using
existing available software used in previous studies\fixme{cite
  previous studies} (we believe there is value in re-creating results
with alternative software).  Our simulations take place within a
three-dimensional Cartesian grid that has a grid spacing of
.05\micron, and our cell shapes are specified as the zero of an
analytic function which can generate the geometry presented in this
paper.  Our code is publically available on github.com~\fixme{cite
  github account in references? Yes!}

@@

\blue{(10)}  Huang \emph{et al.}~\cite{huang2003dynamic} has performed a linear
stability analysis on a cylindrical model which shows an upper limit
on a steady state solution of a $2\micron$ half wavelength.  Cells
with dimensions longer than this spontaneously develop spatial
oscillations, while cells that are shorter relax into a homogeneous
steady state.  We have performed a similar linear stability analysis
on an infinite slab with a thickness equal to that of our flattened
cells ($0.40\micron$).  We first solve the model's five differential
equations for the steady state solution, under the constraint that the
total number MinD and MinE molecules matches our simulation.  We write
down as a matrix the linear response of the time derivatives of the
density to small, spatially harmonic density perturbations around the
steady-state density.  The system is stable at this wavelength
provided all eigenvalues of the matrix are negative, indicating that
all perturbations with this wavelength will decay.  We find the
stability limit by searching for the largest wavelength at which the
system is stable.  Through this method we arrive at an upper half
wavelength stability limit of $1.56\micron$ for our $0.40\micron$
flattened cells. As expected, when decreasing the lengths and widths
of our simulated flattened cells so that the longest distance across
the cell is less than this length, the cells stop exhibiting any
oscillatory behavior.  The deterministic model relaxes into a
motionless state and the stochastic model exhibits random fluctuations
without spatial oscillations.

@@

\blue{(5a),(7)} Both Fig.~\ref{image-p} and its associated movies (see
supplementary material), show that the stochastic model exhibits a
``starry night'' effect characterized by spatially fixed points of
protein density build up, as recruitment leads to clusters of MinD
forming on the wall and then subsequently dissipating.  We attribute
this effect to the fact that the Huang model does not include
diffusion along the membrane, which has been experimentally verified
yet found to be two orders of magnitude less than diffusion in the
cytoplasm~\cite{meacci2005min}. Even ignoring this effect, however, it
is clear from Fig.~\ref{image-p} that the stochastic simulation
results in a protein distribution that is less spatially regular than
the deterministic model predicts.  This naturally leads to
irregularity in the temporal periodicity of the system, which we
quantify by examining the temporal correlation function of the total
MinD found in two opposite polar regions.
%
This correlation function, which is displayed in Fig.~\ref{corr-pill},
is given by
\begin{equation}
  C(\tau) \propto \int
  (N_{\textit{top}}(t) - \bar N_{\textit{top}})
  (N_{\textit{bottom}}(t+\tau) - \bar N_{\textit{bottom}})dt
\end{equation}
where $N_{\textit{top}}(t)$ and $N_{\textit{bottom}}(t)$ are the total
MinD proteins in the top and bottom thirds of a cell.  These plots
help us in studying the periods and the regularity and stability of
oscillations.  These curves are well fit with a simple decoherence
model, given by the equation
\begin{equation}
  C(\tau) = -\cos\left(\frac{2\pi\tau}{T}\right) e^{-\frac{\tau}{\tau_c}}
\end{equation}
where $T$ is the period and $\tau_c$ is the coherence time.  In the
case of the stochastic model the pill exhibits coherence times of
about 8 periods and in the case of the deterministic model oscillate
indefinitely with complete coherence, indicating that its behavior is
perfectly periodic.

\begin{figure*}
  \centering
  \hspace{-3cm}
  \vspace{1cm}
  \includegraphics[width=16cm]{../paper/plot-ave}
  \caption{We display here arrows depicting successive maxima in space
    and and time overlayed on a color plot of the total MinD density
    averaged over the same time period.  The simulation time covered
    for the flattened cells is 350 seconds, which is the same
    period of time depicted in the experimental data plots of Mannik
    \emph{et al.}~\cite{mannik2012robustness}.  For the wild-type pill
    shape, we only cover 250 seconds, in order to provide a
    useful comparison due to its shorter oscillation period.  The top
    row shows plots published by Mannik \emph{et al.} of the MinD maxima behavior
    and the bottom two rows show our simulations using the stochastic
    and deterministic models, respectively.  We simulated
    approximations to the two shapes observed by Mannik, which we call
    \emph{shape A} and \emph{shape B}.  In addition, we studied two
    flattened stadium shapes which we call \emph{stadium A} and
    \emph{stadium B} corresponding in aspect ratio and thickness to
    the two experimental shapes.  The spatial length scale of all the
    figures shown was identical. Each of the flattened cell shapes
    uses the same color scale for the number of proteins per unit
    area.  Finally, we display the natural pill shape, which was also
    featured in Figs.~\ref{image-p} and~\ref{corr-pill}, with a
    different color scale to reflect the thicker cell containing more
    proteins per cross-sectional area.  }
  \label{randst-plot-ave}
\end{figure*}

@@

In every case, including the wild-type pill shape, we see irregularity
in the location of the maxima when using the stochastic model.  The
deterministic model shows uniformly bipolar oscillation. \blue{(2)}
It's worth pointing out that this is a significant result, since for
irregular shapes it would be reasonable to hypothesize a chaotic
system, which would be shown by irregular determinisitic data.  From
these results and how they compare with experiment, we conclude that
the deterministic model is inadequate to explain experimental
observations of the locations of density maxima of the MinD protein.

@@

In contrast to the stochastic model, the original deterministic
version of the model~\cite{huang2003dynamic} predicts behavior that
contradicts experimental observations.  Specifically, the
deterministic model predicts robust and periodic bipolar oscillation
in irregularly shaped cells.  Thus we conclude that this method is
inconsistent with experiment, and cannot be relied upon for
predictions of the behavior of the MinD system.  This is the most
clear example of the inadequacy of the deterministic simulation method
in reproducing experimental behavior of the MinD system to date.
Until now the results of deterministic and stochastic simulation have
largely coincided, with the stochastic method showing minor
differences in behavior when compared to deterministic
models~\cite{kerr2006division, fange2006noise, huang2004min,
  kruse2007experimentalist}.
